An efficient acyclic contact planner for multiped robots

### 2016-01-26 body/appendix_algo.tex
conclusion and appendices as discussed with andrea

\begin{algorithm}
\caption{Discretization of a path} \label{interpolate}
	\begin{algorithmic}[1]
	%~ \Function{GenerateConfiguration}{}
	\Function{Interpolate}{$s0$,$\mathbf{Q}^0$, $MAX\_TRIES$}
		\State $list<$State$>$ $states = [s0]$
		\State $nb\_fail = 0$ 
		\State $i = 1;$ /*Current index in the list*/
		\While {$i < length(\mathbf{Q}^0)$}
			\State State $previous = last\_element(states)$
			\State State $s = \pi(previous, \mathbf{Q}^0[i])$
			\If {$s != NULL $}
				\State $nb\_fail = 0$
				\State $i += 1$
				\State \textbf{return} $\mathbf{q}^{0}$
			\Else
				\State $nb\_fail += 1$
				\If {$nb\_fail == MAX\_TRIES$}
					\State \textbf{return} $FAILURE$
				\EndIf				
				\State $s = $\textsc{RepositionContacts}$(previous)$
			\EndIf
			\State $push\_back(states, s)$
		\EndWhile
		\State \textbf{return} $states$
	\EndFunction
\end{algorithmic}
\label{alg:interpolate}
\end{algorithm}

At each step, $g$ is called with the previous state as a parameter, as well
as a new root placement. $g$ returns a new contact configuration, if it succeeded
in computing a configuration with only one contact switch occurring.
If $g$ failed in achieving this, the method \textsc{RepositionContacts} is called.
It repositions one end effector (either a free limb, or the oldest active contact) towards a new contact position if possible.
This repositioning allows to increase the odds that the contact can be maintained at the next step.

The pseudo code for the method $g$ is given by Algorithm~\ref{alg:pi}.

\begin{algorithm}
\caption{Full body contact generation method} \label{interpolate}
	\begin{algorithmic}[1]
	%~ \Function{GenerateConfiguration}{}
	\Function{$g$}{$pState$,$\mathbf{q}^0$}
		\State State $newState$
		\State $newState.q0 = \mathbf{q}^0$
		\State $newState.freeLimbs = pState.freeLimbs$
		\State /*First try to maintain previous contacts*/
		\State $nbContactsBroken = 0$
		\For {\textbf{each} Limb $k$ in $previous.contactLimbs$}
			\If {$!$\textsc{MaintainContact}$(pState,\mathbf{q}^0,k)$}
				\State $nbContactsBroken += 1$
				\If {$nbContactsBroken > 1$}				
					\State \textbf{return} $NULL$
				\EndIf				
				\State $push(newState.freeLimbs,k)$
			\Else 					
				\State $push(newState.contactLimbs,k)$
			\EndIf
		\EndFor
		\For {\textbf{each} Limb $k$ in $previous.freeLimbs$}
			\If {\textsc{GenerateContact}$(\mathbf{q}^0,k)$}	
				\State $push(newState.contactLimbs,k)$			
				\State \textbf{return} $newState$
			\EndIf
		\EndFor
		\State \textbf{return} $NULL$
	\EndFunction
\end{algorithmic}
\label{alg:pi}
\end{algorithm}

@@

>>>>

\begin{algorithm}
\caption{Discretization of a path} \label{interpolate}
	\begin{algorithmic}[1]
	%~ \Function{GenerateConfiguration}{}
	\Function{Interpolate}{$s0$,$\mathbf{Q}^0$, $MAX\_TRIES$}
		\State $list<$State$>$ $states = [s0]$
		\State $nb\_fail = 0$ 
		\State $i = 1;$ /*Current index in the list*/
		\While {$i < length(\mathbf{Q}^0)$}
			\State State $pState = last\_element(states)$
			\State State $s =$ \textsc{GenFullBody}$(pState, \mathbf{Q}^0[i])$
			\If {$s != NULL $}
				\State $nb\_fail = 0$
				\State $i += 1$
				\State \textbf{return} $\mathbf{q}^{0}$
			\Else
				\State $nb\_fail += 1$
				\If {$nb\_fail == MAX\_TRIES$}
					\State \textbf{return} $FAILURE$
				\EndIf				
				\State $s = $\textsc{RepositionContacts}$(pState)$
			\EndIf
			\State $push\_back(states, s)$
		\EndWhile
		\State \textbf{return} $states$
	\EndFunction
\end{algorithmic}
\label{alg:interpolate}
\end{algorithm}

At each step, \textsc{GenFullBody} is called with the previous state as a parameter, as well
as a new root configuration. \textsc{GenFullBody} returns a new contact configuration, if it succeeded
in computing a configuration with only one contact switch occurring.
If \textsc{GenFullBody} failed in achieving this, the method \textsc{RepositionContacts} is called.
It repositions one end effector (either a free limb, or the oldest active contact) towards a new contact position if possible.
This repositioning allows to increase the odds that the contact can be maintained at the next step.

The pseudo code for the method \textsc{GenFullBody} is given by Algorithm~\ref{alg:pi}.

\begin{algorithm}
\caption{Full body contact generation method} \label{interpolate}
	\begin{algorithmic}[1]
	%~ \Function{GenerateConfiguration}{}
	\Function{\textsc{GenFullBody}}{$pState$,$\mathbf{q}^0$}
		\State State $newState$
		\State $newState.q0 = \mathbf{q}^0$
		\State $newState.freeLimbs = pState.freeLimbs$
		\State /*First try to maintain previous contacts*/
		\State $nbContactsBroken = 0$
		\For {\textbf{each} Limb $k$ in $pState.contactLimbs$}
			\If {$!$\textsc{MaintainContact}$(pState,\mathbf{q}^0,k)$}
				\State $nbContactsBroken += 1$
				\If {$nbContactsBroken > 1$}				
					\State \textbf{return} $NULL$
				\EndIf				
				\State $push(newState.freeLimbs,k)$
			\Else 					
				\State $push(newState.contactLimbs,k)$
			\EndIf
		\EndFor
		\For {\textbf{each} Limb $k$ in $pState.freeLimbs$}
			\If {\textsc{GenerateContact}$(\mathbf{q}^0,k)$}	
				\State $push(newState.contactLimbs,k)$
				\State $remove(newState.freeLimbs,k)$		
				\State \textbf{return} $newState$
			\EndIf
		\EndFor
		\If {\textsc{IsInStaticEquilibrium}$(newState)$}
			\State \textbf{return} $newState$
		\Else
			\State \textbf{return} $NULL$
		\EndIf
	\EndFunction
\end{algorithmic}
\label{alg:pi}
\end{algorithm}

@@

The method \textsc{IsInStaticEquilibrium} returns whether a given state is in static equilibrium.

### 2016-01-26 body/appendix_roms.tex
conclusion and appendices as discussed with andrea

Figure~\ref{fig:hrp2_roms} illustrate the obtained $W^k$ for HRP-2.
Regarding the procedure, we can see that step 2 is conservative (Figure~\ref{fig:hrp2_roms}--right), which 
is acceptable, especially because the lost set essentially relates to configurations close to singularity (they are close to the boundaries of the reachable workspace). We choose again to be less complete
but more efficient, regarding the number of collision tests to be performed by RB-RRT.
In step 1 on the other hand, selecting the convex hull (Figure~\ref{fig:hrp2_roms}--middle) instead of a minimum encompassing shape (Figure~\ref{fig:hrp2_roms}--left) may introduce false positives.
Concretely, because the false positive set intersects with $W^0$, the scaling volume of the robot torso, the induced error is compensated,
as verified by the results shown by Table~\ref{tab:requestpercent}.

>>>>

Figure~\ref{fig:hrp2_roms} illustrate the obtained $W^k$ for HRP-2.
Regarding the procedure, we can see that step 2 is conservative (Figure~\ref{fig:hrp2_roms}--right), which 
is acceptable, especially because the lost set essentially relates to configurations close to singularity (they are close to the boundaries of the reachable workspace, and
often not \gls{contact feasible}, as illustrated in Figure~\ref{fig:dedefeas}, where the exterior boundaries of the reachable workspace appear
red, thus not belonging to \gls{ccontact0}). We choose again to be less complete but more efficient, regarding the number of collision tests to be performed by RB-RRT.
In step 1 on the other hand, selecting the convex hull (Figure~\ref{fig:hrp2_roms}--middle) instead of a minimum encompassing shape (Figure~\ref{fig:hrp2_roms}--left) may introduce false positives.
Concretely, because the false positive set intersects with $W^0$, the scaling volume of the robot torso, the induced error is compensated,
as verified by the results shown by Table~\ref{tab:requestpercent}.

### 2016-01-26 body/conclusion.tex
conclusion and appendices as discussed with andrea

Our results demonstrate that our method allows a pragmatic compromise between three 
criteria that are hard to conciliate: generality, performance, and quality of the solution, making it the first acyclic contact
planner compatible with \gls{interactive} applications.
%
\textbf{Regarding generality}, the \gls{reachc}, coupled with an approach based on limb decomposition, 
allows the method to address arbitrary multiped robots in \gls{cluttered} problems. The only pre-requisite is the specification 
of the volumes $W^0$.
%
\textbf{Regarding performance}, our framework is really efficient in addressing both \Pa and $\mathcal{P}_2$. This results in \gls{interactive} computation times.
%
\textbf{Regarding the quality of the paths}, the \gls{reachc} allows us to compute
\gls{equilibrium feasible} paths in all the presented scenarios, with low rejection rates.
As for \cite{Bouyarmane2009}, failures can still occur, due to the approximate condition used to compute the guide path.
The low computational burden of our framework however allows for fast re-planning in case of failure.

One direction for future work is to focus on a more accurate formulation of $C_{reach}^0$ to address 
this limitation.
Our main objective is of course to generate the complete interpolation between the contact sequences.
We have already obtained some success for some of the computed sequences~\citep{Carpentier2016}, but additional
work is required on the planning side to obtain a seamless workflow. To achieve this, we are currently working on the notion of transition certificate, i.e. formulating
conditions that guarantee that the interpolation between two contact configurations is dynamically feasible.
%~ Our current direction is to propose a geometric quantification of both the kinematic and dynamic constraints of the problem, expressed
%~ at the center of mass of the robot.
 %~ We are also
%~ working on new heuristics more closely related to the dynamics of the systems (for instance, choose the contacts that minimize the jerk between two contact configurations).
Lastly, we aim at going beyond static equilibrium and performing kinodynamic planning. We believe that the most promising direction in this regard is to integrate
the notion of Admissible Velocity Propagation in our current work \citep{DBLP:conf/rss/PhamCN13}.
Addressing these two issues is essential to bridge the gap between the planning and control aspects of multiped locomotion.

>>>>

Our results demonstrate that our method allows a pragmatic compromise between three 
criteria that are hard to conciliate: generality, performance, and quality of the solution, making it the first acyclic contact
planner compatible with \gls{interactive} applications.
%
\textbf{Regarding generality}, the \gls{reachc}, coupled with an approach based on limb decomposition, 
allows the method to address arbitrary multiped robots in \gls{cluttered} problems. The only pre-requisite is the specification 
of the volumes $W^0$.
%
\textbf{Regarding performance}, our framework is really efficient in addressing both \Pa and $\mathcal{P}_2$. This results in \gls{interactive} computation times.
%
\textbf{Regarding the quality of the paths}, the \gls{reachc} allows us to compute
\gls{equilibrium feasible} paths in all the presented scenarios, with low rejection rates.
As for \cite{Bouyarmane2009}, failures can still occur, due to the approximate condition used to compute the guide path.
The low computational burden of our framework however allows for fast re-planning in case of failure.
Furthermore because of this approximation, the guide search is not complete. The choice is deliberate, because we are convinced
that it is necessary to trade completeness for efficiency at all stages of the planner.
However, one direction for future work is to focus on a more accurate formulation of $C_{reach}^0$ to improve
the approximation.

Another limitation of the method is that it currently only applies to \gls{cluttered} problems.
However, it should first be noted that many problems belong to this class: for instance all the problems
at the DRC were \gls{cluttered}. Our method is thus already useful for many cases.
One way of extending its range of application, that we 
consider for future work, is to include the equilibrium criterion when solving $\mathcal{P}_1$.
Considering the set of obstacles intersecting with the reachable workspace for a given root configuration as candidates surfaces, we can use them to verify the equilibrium criterion.
This would give us a necessary condition for \equilibriumfeasibility.

Regarding the interpolation between the contact sequences ($\mathcal{P}_3$), we have already obtained some success for some of the computed sequences~\citep{Carpentier2016}, but additional
work is required on the planning side to obtain a seamless workflow. To achieve this, we are currently working on the notion of transition certificate, i.e. formulating
conditions that guarantee that the interpolation between two contact configurations is dynamically feasible.
%~ Our current direction is to propose a geometric quantification of both the kinematic and dynamic constraints of the problem, expressed
%~ at the center of mass of the robot.
 %~ We are also
%~ working on new heuristics more closely related to the dynamics of the systems (for instance, choose the contacts that minimize the jerk between two contact configurations).
A last limitation of our method is that only static equilibrium configurations are considered for contact planning.
We aim at performing kinodynamic planning to overcome this limitation. We believe that the most promising direction in this regard is to integrate
the notion of Admissible Velocity Propagation in our current work \citep{DBLP:conf/rss/PhamCN13}.
Addressing these two last issues is essential to bridge the gap between the planning and control aspects of multiped locomotion.
